\begin{thebibliography}{999}

\bibitem{sqrt5first}
J.~Bober, A.~Deines, A.~Klages-Mundt, B.~LeVeque, R.~A. Ohana, A.~Rabindranath,
  P.~Sharaba, and W.~Stein, \emph{A {D}atabase of {E}lliptic {C}urves over
  {$\mathbf{Q}(\sqrt{5})$}---{F}irst {R}eport}, Proceedings of the 10th
  International Symposium (ANTS-X) (2012).

\bibitem{BrumerKramer}
A.~Brumer and K.~Kramer, \emph{Paramodular abelian verieties of odd conductor},
arXiv:1004.4699.


\bibitem{cremona1}
J.~E.~Cremona, \emph{Hyperbolic tessellations, modular symbols, and elliptic curves over complex quadratic fields},
Compositio Math. \textbf{51} (1984), no.~3, 275--324.

\bibitem{cremona-lingham}
J.\thinspace{}E. Cremona and M.\thinspace{}P. Lingham, \emph{Finding all
  elliptic curves with good reduction outside a given set of primes},
  Experiment. Math. \textbf{16} (2007), no.~3, 303--312. 

\bibitem{cremona}
J. E. Cremona and E. Whitley, {Periods of cusp forms and elliptic curves over imaginary quadratic fields},
Math. Comp. {\bf 62} (1994), no. 205, 407--429.

%\bibitem{fite-kedlaya-rotger-sutherland}
%F. Fit\'e, K.S. Kedlaya, V. Rotger, and A.V. Sutherland, Sato-Tate distributions and Galois endomorphism
%modules in genus 2, \textit{Compositio Math.} (published online 2012).


\bibitem{quartic}
P.~E. Gunnells, F.~Hajir, and D.~Yasaki, \emph{Modular forms and elliptic
  curves over the field of fifth roots of unity}, to appear in Exp.~Math.

\bibitem{neg23}
  P.~E. Gunnells and A.~Klages-Mundt, \emph{Modular elliptic curves over the
  cubic field of discriminant $-23$}, In preparation.

\bibitem{complexcubic}
 P.~E. Gunnells and D.~Yasaki, 
 \emph{Modular forms and elliptic curves over the
    cubic field of discriminant $-23$}, Int. J. Number Theory, \textbf{9},
 (2013), no.~1, 53--76.

\bibitem{KS-1} K.S.~Kedlaya and A.V.~Sutherland,
Hyperelliptic curves, $L$-polynomials, and random matrices, in
\textit{Arithmetic, Geometry, Cryptography, and Coding Theory (AGC$^2$T
2007)}, Contemp. Math. 487, Amer. Math. Soc., 2009.

\bibitem{kedlaya-sutherland-ants}
K.S.~Kedlaya and A.V.~Sutherland,
{\em Computing $L$-series of hyperelliptic curves},
in Algorithmic Number Theory (ANTS VIII),
Lecture Notes in Comp. Sci. 5011, Springer, 2008, 312--326.

\bibitem{poory}
C.~{Poor} and D.~S. {Yuen}, 
\emph{Computations of spaces of Siegel modular cusp forms},
J. Math. Soc. Japan \textbf{59} (2007), no.~1, 185--222. 

\bibitem{PY_para_p}
C.~{Poor} and D.~S. {Yuen}, 
\emph{{Paramodular Cusp Forms}}, ArXiv e-prints
  (2009).


\end{thebibliography}
