\documentclass[12pt]{article}
%\usepackage{charter}

\setlength{\textwidth}{6in}
\setlength{\oddsidemargin}{0.25in}
\setlength{\topmargin}{0in}
\setlength{\textheight}{8in}
\renewcommand{\baselinestretch}{1.4}

%\usepackage{amsfonts}
\usepackage{url}
\usepackage{verbatim}

\renewcommand{\refname}{\centerline{\bf \large References Cited}}

\newcommand{\nc}{\newcommand}
\nc{\noi}{\noindent}
\nc{\cmmt}[1]{}

\newcommand{\secti}[1]{\par\medskip\centerline{\textsc{#1}}\par\medskip}

\begin{document}
\begin{center}
\textbf{\large  Project Summary}
\end{center}
\bigskip

We request funding for a conference in the number theory on connections
between curves, modular forms, $L$-functions, and computation to be
held at Arizona State University.  Some of the top number theorists in
these areas have agreed to participate, and Arizona State University
has agreed to provide infrastructural support.


\medskip
\noindent\textbf{Intellectual Merit:}

\medskip

The conference will present lectures on recent developments in number theory
in the mornings, and
bring together experts on related topics to encourage collaborative
research in the afternoons in a workshop atmosphere.  Areas of focus for 
the conference are hyperelliptic curves over the rationals, elliptic
curves over low degree number fields, Hilbert modular forms, Siegel
modular forms, and in all cases, corresponding $L$-functions.  These
are broad topics; we will pay particular attention to computationally
related questions.

\medskip
\noindent\textbf{Broader Impact:}

\medskip

We are committed to having participants from all career stages, as well
as having substantial participation by mathematicians from underrepresented
groups.  For graduate students, postdoctoral fellows, and junior faculty,
the confernce can serve as a springboard for their future in the profession.

Computational results presented or worked on at the conference will be
encorporated into the $L$-functions and Modular Forms Database, an
publicly available web site which is the result of an international
collaboration of many number theorists.  The site provides data which is
useful for mathematical research, and serves as an educational tool
helping both young and experienced number theorists see some of the
objects and subtle relationships which are an important part of modern
number theory.


\end{document}
